Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-09-12 paper/paper.tex

\section{Benchmarks and case studies}
\subsection{Performance benchmarks}
We run a single thread benchmark on CPU Intel(R) Xeon(R) CPU E5-2686 v4 @ 2.30GHz,
and its CUDA version on a GPU Tesla V100.
The results are shown in Figure~\ref{fig:benchmark}.
The graph we use are random 3 regular graphs, a typical type of sparse graph that having small tree width $\text{tw}(G) \leq |V|/6$~\cite{Fomin2006}.

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets on a random three regular graph with different tensor element types.
    (b) is the computing time for computing maximum independent set size, number of independent sets and suboptimal sets.
    (c) is the computing time for independence polynomials with different approaches.
    (d) is the computing time for finding all configurations and a single configuration, with or without bounding.
    }
    \label{fig:benchmark}
\end{figure}

Among the benchmarks shown in \Fig{fig:benchmark}, computing \texttt{max size (CPU)},
\texttt{total size (CPU)}, \texttt{total size (GPU)}, \texttt{IDP (FFT) (CPU)}, \texttt{IDP (FFT) (GPU)} and \texttt{single configuration (bounding) (CPU)} can utilize fast BLAS functions,
hence they are relatively faster comparing to other methods in the same category.
Only algebras requiring dynamic sized structures can not be computed on GPU, here are polynomial algebra for computing independence polynomial and set algebra for enumerating configurations.
When computing the independence polynomial, the finite field approach is the only method without round off error that can run on GPU.
To enumerate the MISs, the technique to bound the enumeration space is important,
without which the largest enumeratible size is only $150$ with 32G memory while being one order slower.
All these computing times have strong correlation with the tree width,
while the computing times of configurations enumeration also depends on the size of target configuration space.

>>>>

\section{Benchmarks and case studies}
\subsection{Performance benchmarks}
We run a single thread benchmark on CPU Intel(R) Xeon(R) CPU E5-2686 v4 @ 2.30GHz,
and its CUDA version on a GPU Tesla V100.
The results are summarized in Figure~\ref{fig:benchmark}.
The type of graphs we used for benchmarks here are random 3 regular graphs, a typical type of sparse graphs that has a small tree width $\text{tw}(G) \leq |V|/6$~\cite{Fomin2006}.

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets on a random three regular graph with different tensor element types. \red{for each graph size, is there only one graph?} (a) treewidth versus the number of vertices for the benchmarked graphs. 
    (b) The computing time for calculating the independence number and for counting of the number of independent sets, the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$.
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for enumeration problems, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    \red{suggestions for the figure: 1. the legend seems to overlap the lines, especially the large subfigure. Not sure what's a better way to include the legend there. 2. I see you guys don't typically use markers. We typically use markers to distinguish the different lines, as well as using the color. Especially, for some journals, there is a physical non-color copy, so some requires the figures to be understandable without colors.}
    }
    \label{fig:benchmark}
\end{figure}

The computations shown in \Fig{fig:benchmark} include \texttt{max size} (independence number, $\alpha(G)$), \red{we can consider changing `max size' to `ind number', but I am okay with both}
 \texttt{total counting} (total number of independent sets) \red{should we change total size to total counting here, since max size meant max size of the set}, 
 \texttt{max counting} (counting of MISs), 
 \texttt{max2 counting} (counting of independent sets having a size $\alpha(G)$ or $\alpha(G)-1$), \red{we can consider changing `max counting' and `max2 counting' to `MIS counting' and `MIS and MIS-1 counting', but I am okay with both notations}
 \texttt{IDP} (independence polynomial), 
 \texttt{single MIS} (find one MIS configuration), 
 \texttt{single MIS (bounding)} (find one MIS configuration including the bounding of enumeration space), \texttt{all MIS} (enumerate all MIS configurations), 
\texttt{all MIS (bounding)} (enumerate all MIS configurations including the bounding of enumeration space),
\texttt{all MIS and MIS-1} (enumeration of all independent sets having a size $\alpha(G)$ or $\alpha(G)-1$),
\texttt{all IS} (enumeration of all independent sets).
\red{Jinguo, should we include maximal independence polynomial and enumeration of all maximal independent sets as well? I think that will make the benchmark more complete.}

Among the benchmarks, computing \texttt{max size (CPU)},
\texttt{total size (CPU)}, \texttt{total size (GPU)}, \texttt{IDP (FFT) (CPU)}, \texttt{IDP (FFT) (GPU)} and \texttt{single configuration (bounding) (CPU)} can utilize fast BLAS functions, and hence 
they are faster comparing to non-BLAS methods in the same category.
Only algebras requiring dynamic sized structures cannot be computed on GPU, which includes the polynomial algebra for computing the independence polynomial and the set algebra for enumerating configurations.
When computing the independence polynomial, the finite field approach is the only method that does not incur round-off errors and that can be run on a GPU.
To enumerate the MISs, the technique to bound the enumeration space is indispensible,
without which the largest enumerable size is about $150$ vertices with 32G memory while also being one order-of-magnitude slower.
All these computing times have a strong correlation with the treewidth, 
and the computing times of configuration enumeration also depend on the size of the target configuration space.
